---
title: "ESM 244 Lab 5"
author: "Tom Paschos"
date: "2/6/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

a. Attach packages

```{r Load_Packages_Data}

# tidyverse, tseries, forecast

library(tidyverse)
library(tseries)
library(forecast)

```

b. Get data

US Residential energy consumption values (Jan 1973 - Oct 2017), Trillion BTU (US EIA)

```{r}

energy <- read_csv("energy.csv")

```

c. Create 'ts' time series data using info in the res_total column from 'energy'

```{r}

res_ts <- ts(energy$res_total, frequency = 12, start = c(1973,1))

res_ts

plot(res_ts) # plot function in R is awesome to explore

# Data looks trending, additive, but there are also intermediate lower peaks that get incraesgingly large, so let's decompose

```

d. Decompose ts

```{r}

res_dc <- decompose(res_ts)

plot(res_dc) # this does a ton of work for us - 1st graph: original data; 2nd graph: trend; 3rd graph: seasonality; 4th graph: residuals

```

e. Other cool ways to present ts data

```{r}

monthplot(res_ts)

```

```{r}

ggseasonplot(res_ts, polar = TRUE) # This displays ts data with seasonality exploration

# polar shows seasonality

```

f. Simple moving averages

forecast::ma() allows you to manually specify what the window size to be

```{r}

sma_res <- ma(res_ts, order = 5) # Changing the order (or window, or moving average) can drastically alter how the data appears.  Best option is to show the original data AND the moving average

plot(res_ts)
lines(sma_res, col = "red")

```

g. ACF (autocorrelation)

```{r}

res_acf <- acf(res_ts) # This is just in the base stats R package

# You can see, as expected, these drastic peaks every 12 months.  Things are very highly correlated with obs that existed 12 months before then.

#ggtsdisplay() can show more info

ggtsdisplay(res_ts)

```

i. Holt-Winters exponential smoothing 

```{r}

res_hw <- HoltWinters(res_ts)
res_hw

plot(res_hw)

```

Now, some future forecasting with Holt-Winters

```{r}

res_forecast <- forecast(res_hw, h = 60)
plot(res_forecast)

```

Check distribution of the residuals:

```{r}

hist(res_forecast$residuals)
# could aslo do a qqplot
qqnorm(res_forecast$residuals)

```

ARIMA - Auto Regressive Integrated Moving Average.  Using auto.arima for p,d,q

```{r}

res_pdq <- auto.arima(res_ts)
res_pdq

# Results: ARIMA (1,0,2)(0,1,1)[12] with drift

```

pdq: (1,0,2), (0,1,1)

Fit the ARIMA model using these pdq inputs:

```{r}

res_arima <- arima(res_ts, order = c(1,0,2), seasonal = list(order = c(0,1,1))) # first non-seasonal pdq, then the seasonal one
res_arima

# Check out the residuals

par(mfrow = c(1,2)) # this says anything below this populate in 1 row with 2 columnes
hist(res_arima$residuals)
qqnorm(res_arima$residuals)

# Nothing major stands out

```

Now forecast with ARIMA outputs:

```{r}

forecast_res <- forecast(res_arima, h = 72) # forecasted to 72 months
plot(forecast_res)

```

Now let's try viewing this in ggplot:

```{r}

res_df <- data.frame(forecast_res)
month_seq <- seq(1,72)

res_df_2 <- data.frame(month_seq, res_df)

ggplot(res_df_2, aes(x = month_seq, y = Point.Forecast)) +
  geom_line() # Could add CI ribbons if you wanted

```

End of Lab
